---
title: "VAR analysis - Granger-Causality-Initial Comparisons"
author: "Katie Schilling - 501130072"
date: "07/02/2022"
output:
  pdf_document: default
  html_document: default
---
Required tools to be loaded
```{r}
library(dlookr)
library(lubridate)
library(dplyr)
library(ggpubr)
library(forecast)
library(funModeling)
require(ggfortify)
require(tseries)
require(MTS)
require(vars)
require(fUnitRoots)
require(lattice)

```

Import both Clean datasets
```{r}
Covid_monthly <- read.csv("C:/Users/Katie Schilling/Downloads/covid_monthly_clean.csv")
Vital_Events <- read.csv("C:/Users/Katie Schilling/Downloads/vital_events_clean.csv")

```

Combine the vital events data with the Covid Monthly data
```{r}
Final_dataset <- merge(x=Vital_Events, y=Covid_monthly, all = TRUE)
```

Check the data and ensure data merged properly
```{r}
summary(Final_dataset)
```

Change the N/A in the Covid Positive Cases to 0 so that the data is not omitted from the predictions
```{r}
Final_dataset$Covid[is.na(Final_dataset$Covid)] = 0
```

See if there are anymore NA's in the dataset
```{r}
Final_dataset %>% filter_all(any_vars(is.na(.))) 
```
Remove rows with NA as they will skew the results
```{r}
Final_dataset <- na.omit(Final_dataset)
```

Check for NA's to confirm all have been removed
```{r}
Final_dataset %>% filter_all(any_vars(is.na(.))) 
```
```{r}
summary(Final_dataset)
```

```{r}
Final_dataset$Date <- as.Date(Final_dataset$Date,"%Y-%m-%d")

```


```{r}
glimpse(Final_dataset)
```


Normalize the Data
```{r}

Final_dataset_standardized <- Final_dataset %>% mutate_each_(list(~scale(.) %>% as.vector),
                                  vars = c("Births","Marriages","Deaths","Stillbirths","Covid"))

```


Convert data frame to a time series
```{r}
Final_dataset_TS <- ts(Final_dataset_standardized[2:6], frequency = 12, start = 1994, end = 2021)
```

```{r}
plot(Final_dataset_TS[,1], ylab="", yaxt="n", col = "red", lwd = 2)
title(main = "Births Timeseries", xlab = "Time", ylab = "Births")

plot(Final_dataset_TS[,3], ylab="", yaxt="n", col = "red", lwd = 2)
title(main = "Deaths Timeseries", xlab = "Time", ylab = "Deaths")

plot(Final_dataset_TS[,2], ylab="", yaxt="n", col = "red", lwd = 2)
title(main = "Marriages Timeseries", xlab = "Time", ylab = "Marriages")

plot(Final_dataset_TS[,4], ylab="", yaxt="n", col = "red", lwd = 2)
title(main = "Stillbirths Timeseries", xlab = "Time", ylab = "Stillbirths")

plot(Final_dataset_TS[,5], ylab="", yaxt="n", col = "red", lwd = 2)
title(main = "Covid Timeseries", xlab = "Time", ylab = "Covid")
##ts.plot(Final_dataset_TS)
```

```{r}
xyplot.ts(Final_dataset_TS[,1])
```

---------------------------------------------------------------------------------------------------------------------------------------------------

decompose data and display results
```{r}
decomp <- decompose(Final_dataset_TS)
plot(decomp$seasonal)
plot(decomp$trend)
plot(decomp$random)
```


```{r}
Covid <- forecast(Final_dataset_TS[,5])
plot(Covid, main = "Covid Forecast")

```
```{r}
Stillbirths <- forecast(Final_dataset_TS[,4])
plot(Stillbirths, main = "Stillbirths Forecast")

```
```{r}
Deaths <- forecast(Final_dataset_TS[,3])
plot(Deaths, main = "Deaths Forecast")

```
```{r}
Marriages <- forecast(Final_dataset_TS[,2])
plot(Marriages, main = "Marriages Forecast")

```
```{r}
Births <- forecast(Final_dataset_TS[,1])
plot(Births, main = "Births Forecast")

```


---------------------------------------------------------------------------------------------------------------------------------------------------VAR time series forecasting for multivarite

```{r}
apply(Final_dataset_TS, 2, adf.test)
```


```{r}
plot.ts(Final_dataset_TS, main= "Final Dataset Timeseries")
```

```{r}
autoplot(ts(Final_dataset_TS,
            start = c(1964,1),
            frequency = 12)) +
  ggtitle("Time Series Plot of the stationary `Final Dataset' Time-Series")
```

```{r}
VARselect(Final_dataset_TS, 
          type = "none", 
          lag.max = 6) #highest lag order
```

```{r}
# Creating a VAR model with vars
final_ts_var <- vars::VAR(Final_dataset_TS,
                   lag.max = 6, #highest lag order for lag length selection according to the choosen ic
                   ic = "AIC", #information criterion
                   type = "none") #type of deterministic regressors to include
summary(final_ts_var)

```

```{r}
serial.test(final_ts_var)
```

VAR model causality test for each variable against all other variables
```{r}
causality(final_ts_var, 
          cause = c("Deaths"))

```

```{r}
causality(final_ts_var,
          cause = c("Births")) 

```

```{r}
causality(final_ts_var, 
          cause = c("Covid")) 

```

```{r}
causality(final_ts_var, 
          cause = c("Marriages"))  

```

```{r}
causality(final_ts_var,
          cause = c("Stillbirths"))

```
prediction for final dataset including covid
```{r}
fcast = predict(final_ts_var, n.ahead = 12, ci = 0.95) # predicting the 12 months after the data ends
par(mar = c(2,2,2,2))
plot(fcast)
```



forecast for final dataset including covid
```{r}
Final_forecast <- forecast(Final_dataset_TS, h=12 ) 
par(mar = c(2,2,2,2))
plot(Final_forecast)

```


```{r}
accuracy(Final_forecast,d= NULL, D=NULL)

```

Create forecast for vital events excluding Covid to see if the introduction of covid had a large impact on the other variables

```{r}
Vital_Events$Date <- as.Date(Final_dataset$Date,"%Y-%m-%d")

```


```{r}
glimpse(Vital_Events)
```


Normalize the Data
```{r}

Vital_Events_standardized <- Final_dataset %>% mutate_each_(list(~scale(.) %>% as.vector),
                                  vars = c("Births","Marriages","Deaths","Stillbirths"))

```

Convert data frame to a time series up to 2019, so that I can compare predictions for 2021-2021 to actual observed numbers
```{r}
vital_events_TS <- ts(Vital_Events_standardized[2:5], frequency = 12, start = 1994, end = 2019)
```


```{r}
plot(vital_events_TS)
```

```{r}
xyplot.ts(vital_events_TS)
```

---------------------------------------------------------------------------------------------------------------------------------------------------

decompose data and display results vital events without covid
```{r}
decomp <- decompose(vital_events_TS)
plot(decomp$seasonal)
plot(decomp$trend)
plot(decomp$random)
```

Dicky-Fuller test
```{r}
apply(vital_events_TS, 2, adf.test)
```
```{r}
autoplot(ts(vital_events_TS,
            start = c(1994,1),
            frequency = 12)) +
  ggtitle("Time Series Plot of the stationary `Vital Events Dataset' Time-Series")
```

```{r}
VARselect(vital_events_TS, 
          type = "none", lag.max = 6) 
```
Creating a VAR model with vars
```{r}

vital_ts_var <- vars::VAR(vital_events_TS,
                   lag.max = 6, 
                   ic = "AIC", 
                   type = "none") 
summary(vital_ts_var)
```

```{r}
causality(vital_ts_var, 
          cause = c("Deaths"))

```

```{r}

causality(vital_ts_var,
          cause = c("Births")) 

```


```{r}

causality(vital_ts_var, 
          cause = c("Marriages"))  

```

```{r}

causality(vital_ts_var,
          cause = c("Stillbirths")) 

```
prediction of vital events without covid
```{r}

fcast2 = predict(vital_ts_var, n.ahead = 24) # predicting the 12 months after the data ends
par(mar = c(2,2,2,2))
plot(fcast2)

fcast2

fcast3 = predict(vital_ts_var, n.ahead = 12) # predicting the 12 months after the data ends
par(mar = c(2,2,2,2))
plot(fcast3)
fcast3
head(Final_dataset)
```



forecast for just the vital events, without covid
```{r}
vital_forecast <- forecast(vital_ts_var, h=24 ) #Forecasting 12 months ahead
par(mar=c(5.1,4.1,4.1,2.1))
autoplot(vital_forecast, alpha = 0.5)

```
checking accuracy of both forecasts
```{r}
accuracy(vital_forecast,d= NULL, D=NULL)
accuracy(Final_forecast,d= NULL, D=NULL)
```



Create dataframes for both acutal values and predicted values for comparison
This is for vital events only, without the introduction of Covid to see how the predicted compared, and how much
variance there is between the 2 and the impact covid had on the actual numbers
```{r}
library(modelr)
actuals <- as.data.frame(Final_dataset_standardized[1:5])
actuals <- actuals[313:325,]
predicted <- fortify(stats::predict(vital_ts_var, n.ahead = 24))
predicted <- predicted[313:325,]
predicted <- predicted[c(1,6,10,14,18)]
colnames(predicted)<- c("Date","Births","Marriages","Deaths", "Stillbirths")

covid_stand <-  Covid_monthly %>% mutate_each_(list(~scale(.) %>% as.vector),
                                  vars = c("Covid"))
covid_stand$Date <- as.Date(covid_stand$Date,"%Y-%m-%d")
summary(covid_stand)
```


Predicted vs Actuals Births
```{r}
ggplot() +

      geom_point(data=predicted, aes(x=Date, y=Births, color="Predicted")) + 
      geom_line(data=actuals, aes(x=Date, y=Births,color="Actuals"))  +
      geom_line(data=covid_stand, aes(x = Date, y=Covid,color="Covid"))  +
      ylim(-3,+2)+
      labs(color='Legend', title = "Predicted vs. Actual Births + Covid I ntroduction")+
      theme(legend.position = c("bottom"), plot.title = element_text(hjust = 0.5)) +
      scale_x_date(limits = as.Date(c("2020-01-01","2021-01-01")))
```
Predicted vs Actuals Deaths
```{r}
ggplot() +

      geom_point(data=predicted, aes(x=Date, y=Deaths, color="Predicted")) + 
      geom_line(data=actuals, aes(x=Date, y=Deaths,color="Actuals"))  +
      geom_line(data=covid_stand, aes(x = Date, y=Covid,color="Covid"))  +
      ylim(-2,+4)+
      labs(color='Legend', title = "Predicted vs. Actual Deaths + Covid Introduction")+ theme(legend.position = c("bottom"), plot.title = element_text(hjust = 0.5))+
      scale_x_date(limits = as.Date(c("2020-01-01","2021-01-01")))
```


```{r}
ggplot() +

      geom_point(data=predicted, aes(x=Date, y=Marriages, color="Predicted")) + 
      geom_line(data=actuals, aes(x=Date, y=Marriages,color="Actuals"))  +
      geom_line(data=covid_stand, aes(x = Date, y=Covid,color="Covid"))  +
      ylim(-2,+4)+
      labs(color='Legend', title = "Predicted vs. Actual Marriages + Covid Introduction")+ theme(legend.position = c("bottom"), plot.title = element_text(hjust = 0.5))+
      scale_x_date(limits = as.Date(c("2020-01-01","2021-01-01")))
```
```{r}
ggplot() +

      geom_point(data=predicted, aes(x=Date, y=Births, color="Predicted")) + 
      geom_line(data=actuals, aes(x=Date, y=Births,color="Actuals"))  +
      geom_line(data=covid_stand, aes(x = Date, y=Covid,color="Covid"))  +
      ylim(-2,+4)+
      labs(color='Legend', title = "Predicted vs. Actual Births + Covid Introduction")+ theme(legend.position = c("right"), plot.title = element_text(hjust = 0.5))+
      scale_x_date(limits = as.Date(c("2020-01-01","2021-01-01")))
```
```{r}
ggplot() +

      geom_point(data=predicted, aes(x=Date, y=Stillbirths, color="Predicted")) + 
      geom_line(data=actuals, aes(x=Date, y=Stillbirths,color="Actuals"))  +
      geom_line(data=covid_stand, aes(x = Date, y=Covid,color="Covid"))  +
      ylim(-2,+4)+
      labs(color='Legend', title = "Predicted vs. Actual Stillbirths + Covid Introduction")+ theme(legend.position = c("bottom"), plot.title = element_text(hjust = 0.5))+
      scale_x_date(limits = as.Date(c("2020-01-01","2021-01-01")))
```




