---
title: "Combined Data EDA"
author: "Katie Schilling - 501130072"
date: "04/04/2022"
output:
  pdf_document: default
  html_document: default
---


```{r}
library(readxl)
library(caTools)
library(dplyr)
require(lmtest)
library(dlookr)
```

Import bothe Clean datasets
```{r}
Covid_monthly <- read.csv("C:/Users/Katie Schilling/Downloads/covid_monthly_clean.csv")
Vital_Events <- read.csv("C:/Users/Katie Schilling/Downloads/vital_events_clean.csv")

```

Combine the vital events data with the Covid Monthly data
```{r}
Final_dataset <- merge(x=Vital_Events, y=Covid_monthly, all = TRUE)
```

Check the data and ensure data merged properly
```{r}
summary(Final_dataset)
```

Change the N/A in the Covid Positive Cases to 0 so that the data is not omitted from the predictions
```{r}
Final_dataset$Covid[is.na(Final_dataset$Covid)] = 0
```

See if there are anymore NA's in the dataset
```{r}
Final_dataset %>% filter_all(any_vars(is.na(.))) 
```
Remove rows with NA as they will skew the results
```{r}
Final_dataset <- na.omit(Final_dataset)
```

Check for NA's to confirm all have been removed
```{r}
Final_dataset %>% filter_all(any_vars(is.na(.))) 
```
```{r}
summary(Final_dataset)
```

```{r}
Final_dataset$Date <- as.Date(Final_dataset$Date,"%Y-%m-%d")

```


```{r}
glimpse(Final_dataset)
```
Create function to view basic EDA
```{r}
basic_eda <- function(data)
{
  glimpse(data)
  print(status(data))
  freq(data) 
  print(profiling_num(data))
  plot_num(data)
  describe(data)
}  
```



Visual analysis of the normality of the data.  
```{r}
plot_normality(Final_dataset)
```

Find the correlation, if any, between the variables in the data
```{r}
correlate(Final_dataset)
```
visualize the correlation, if any, of the data
```{r}
plot_correlate(Final_dataset)
```

```{r}
hist(Final_dataset$Births)
hist(Final_dataset$Deaths)
hist(Final_dataset$Marriages)
hist(Final_dataset$Stillbirths)
hist(Final_dataset$Covid)
```


```{r}
boxplot(Final_dataset$Births)
boxplot(Final_dataset$Deaths)
boxplot(Final_dataset$Marriages)
boxplot(Final_dataset$Stillbirths)
boxplot(Final_dataset$Covid)

```

```{r}
eda_web_report(Final_dataset)
```