---
title: "EDA - Vital Events"
author: "Katie Schilling - 501130072"
date: "07/02/2022"
output: html_document
---
Required tools to be loaded
```{r}
library(dlookr)
library(lubridate)
library(data.table)
library(dplyr)
library(moments)
library(ggpubr)
library(smooth)
library(greybox)
library(forecast)
library(funModeling)
```

Import the first data set
```{r}
Vital_events<- read.csv("C:/Users/Katie Schilling/Downloads/vital_events_data_by_month_1994-2021_q2 (1).csv", header = TRUE)
## Change the column names to cleaner versions
colnames(Vital_events)<- c("Month","Year","Live Births", "Marriages","Deaths","Stillbirths")
```

View the data to see what information is present
```{r}
head(Vital_events)
```
Clean up of the vitals Data event.  Removal of the french version of the month, as many of them did not import properly.  Makes the data easier to read, view and work with.
```{r}
Vital_events[Vital_events == "January/janvier"] <- "January"
Vital_events[Vital_events == "February/février"] <- "February"
Vital_events[Vital_events == "March/mars"] <- "March"
Vital_events[Vital_events == "April/avril"] <- "April"
Vital_events[Vital_events == "May/mai"] <- "May"
Vital_events[Vital_events == "June/juin"] <- "June"
Vital_events[Vital_events == "July/juillet"] <- "July"
Vital_events[Vital_events == "August/août"] <- "August"
Vital_events[Vital_events == "September/septembre"] <- "September"
Vital_events[Vital_events == "October/octobre"] <- "October"
Vital_events[Vital_events == "November/novembre"] <- "November"
Vital_events[Vital_events == "December/décembre"] <- "December"


## Check data now to see if the changes are sufficent
head(Vital_events)
```

Check for any missing values
```{r}
sum(is.na(Vital_events))
```
Create function to view basic EDA
```{r}
basic_eda <- function(data)
{
  glimpse(data)
  print(status(data))
  freq(data) 
  print(profiling_num(data))
  plot_num(data)
  describe(data)
}  
```


produce and view the results of the EDA function
```{r}
basic_eda(Vital_events)
```

Visual analysis of the normality of the data.  
```{r}
plot_normality(Vital_events )
```

Find the correlation, if any, between the variables in the data
```{r}
correlate(Vital_events)
```
visualize the correlation, if any, of the data
```{r}
plot_correlate(Vital_events)
```

```{r}
hist(Vital_events$`Live Births`)
hist(Vital_events$Deaths)
hist(Vital_events$Marriages)
hist(Vital_events$Stillbirths)
```


```{r}
boxplot(Vital_events$`Live Births`)
boxplot(Vital_events$Deaths)
boxplot(Vital_events$Marriages)
boxplot(Vital_events$Stillbirths)
```

```{r}
eda_web_report(Vital_events)
```