---
title: "Covid Cases Pre-Processing"
author: "Katie Schilling - 501130072"
date: "07/02/2022"
output: html_document
---
Required tools to be loaded
```{r}
library(dlookr)
library(lubridate)
library(data.table)
library(dplyr)
library(moments)
library(ggpubr)
library(smooth)
library(greybox)
library(forecast)
library(funModeling)
```

Import second dataset in regards to COVID-19
```{r}
Covid_positivity <- read.csv("C:/Users/Katie Schilling/Downloads/covid19-download.csv")
```

View the data to see what information is present
```{r}
head(Covid_positivity)
```
Take only the 3 data columns from the data that I want to work with, is province, and NumToday, which is the current numbers for today, and the date.  g the date
```{r}
Covid_positivity <- data.frame(Covid_positivity$date,Covid_positivity$prname, Covid_positivity$numtoday)
```

Assign the date column the proper date format
```{r}
Covid_positivity$Covid_positivity.date <-ymd(Covid_positivity$Covid_positivity.date)
```

remove rows for provinces, and only use Canada as it includes all cases from all provinces.
```{r}
Covid_positivity <- Covid_positivity[Covid_positivity$Covid_positivity.prname == "Canada",]
```

create new colums for both the year and month to allow the data to be aggregated and matched to the vitals event data
```{r}
Covid_positivity$Year <- format(as.Date(Covid_positivity$Covid_positivity.date, "%Y/%m/%d"), "%Y")
Covid_positivity$Month <- format(as.Date(Covid_positivity$Covid_positivity.date, "%Y/%m/%d"), "%m")

Covid_positivity$Covid_positivity.numtoday <- as.numeric(as.character(Covid_positivity$Covid_positivity.numtoday))
```


```{r}
Covid_data_table <- data.table(Covid_positivity)
```

Split each month, for each year into its own dataset so that the data can be combined and I can get a singular number of postie cases for each month.
There was no clear or easy way that I could find to tally up the number of cases for each day into a monthly total, without separating each month into its own data frame.  The days for each month were inconsistent, as some months reports were daily, and other months the reporting was sporadic.
```{r}
February_2020 <- Covid_data_table[Month == "02" & Year == "2020"]
March_2020 <- Covid_data_table[Month == "03" & Year == "2020"]
April_2020 <- Covid_data_table[Month == "04" & Year == "2020"]
May_2020 <- Covid_data_table[Month == "05" & Year == "2020"]
June_2020 <- Covid_data_table[Month == "06" & Year == "2020"]
July_2020 <- Covid_data_table[Month == "07" & Year == "2020"]
August_2020 <- Covid_data_table[Month == "08" & Year == "2020"]
September_2020 <- Covid_data_table[Month == "09" & Year == "2020"]
October_2020 <- Covid_data_table[Month == "10" & Year == "2020"]
November_2020 <- Covid_data_table[Month == "11" & Year == "2020"]
December_2020 <- Covid_data_table[Month == "12" & Year == "2020"]
January_2021 <- Covid_data_table[Month == "01" & Year == "2021"]
February_2021 <- Covid_data_table[Month == "02" & Year == "2021"]
March_2021 <- Covid_data_table[Month == "03" & Year == "2021"]
April_2021 <- Covid_data_table[Month == "04" & Year == "2021"]
May_2021 <- Covid_data_table[Month == "05" & Year == "2021"]
June_2021 <- Covid_data_table[Month == "06" & Year == "2021"]
July_2021 <- Covid_data_table[Month == "07" & Year == "2021"]
August_2021 <- Covid_data_table[Month == "08" & Year == "2021"]
September_2021 <- Covid_data_table[Month == "09" & Year == "2021"]
October_2021 <- Covid_data_table[Month == "10" & Year == "2021"]
November_2021 <- Covid_data_table[Month == "11" & Year == "2021"]
December_2021 <- Covid_data_table[Month == "12" & Year == "2021"]
```

View one of the datasets to see understand and ensure the right information was pulled into the new dataset
```{r}
head(February_2020)
```

Preemptively create new data frame for monthly totals once they are calculated
```{r}
Columns = c("Year", "Month","Covid Positive Cases")
Covid_monthly <- data.frame(matrix(nrow = 1, ncol = length(Columns)))
colnames(Covid_monthly) = Columns
```

Sum columns for numtoday to get a single line of data for each month
Side Note:  This was a challenge.  I was having a hard time finding a simple way to do this, while keeping the data clean and easy to work with.  Once I was able to get the end result I was looking for it was quick to do the same for the rest, but resulted in a lot of lines of code.  I realize this may be a convuluted way to get to the end result, and would appreciate any tips on how I could have streamlined this process for the future!

Each month consists of a chunk of code starting with:
Line 1: Finding the sum of num today to get the total monthly number.
Line 2: Assigning that value to the dataframe.  Unfortunately what I realized is that it over rode all lines of code to have the same number, which is why
Line 3: Only leaves a single line of code, by finding only the rows that are distinct and removing duplicates
Line 4: Change the value in the montly column from the number to the name of the Month to match the vitals even data
Line 5, 6 & 7: Re-naming the columns

```{r}
February = sum(February_2020$Covid_positivity.numtoday)
February_2020 <- data.frame(February_2020$Month, February_2020$Year, February)
February_2020 <- February_2020 %>% distinct(February_2020$Month, .keep_all = TRUE)
colnames(February_2020)[which(names(February_2020) == "February")] <- "Covid Positive Cases"
February_2020["February_2020.Month"][February_2020["February_2020.Month"]== "02"]  <- "February"
colnames(February_2020)[which(names(February_2020) == "February_2020.Year")] <- "Year"
colnames(February_2020)[which(names(February_2020) == "February_2020.Month")] <- "Month"

```

```{r}
March = sum(March_2020$Covid_positivity.numtoday)
March_2020 <- data.frame(March_2020$Month, March_2020$Year, March)
March_2020 <- March_2020 %>% distinct(March_2020$Month, .keep_all = TRUE)
colnames(March_2020)[which(names(March_2020) == "March")] <- "Covid Positive Cases"
March_2020["March_2020.Month"][March_2020["March_2020.Month"]== "03"]  <- "March"
colnames(March_2020)[which(names(March_2020) == "March_2020.Year")] <- "Year"
colnames(March_2020)[which(names(March_2020) == "March_2020.Month")] <- "Month"
```

```{r}
April = sum(April_2020$Covid_positivity.numtoday)
April_2020 <- data.frame(April_2020$Month, April_2020$Year, April)
April_2020 <- April_2020 %>% distinct(April_2020$Month, .keep_all = TRUE)
colnames(April_2020)[which(names(April_2020) == "April")] <- "Covid Positive Cases"
April_2020["April_2020.Month"][April_2020["April_2020.Month"]== "04"]  <- "April"
colnames(April_2020)[which(names(April_2020) == "April_2020.Year")] <- "Year"
colnames(April_2020)[which(names(April_2020) == "April_2020.Month")] <- "Month"
```

```{r}
May = sum(May_2020$Covid_positivity.numtoday)
May_2020 <- data.frame(May_2020$Month, May_2020$Year, May)
May_2020 <- May_2020 %>% distinct(May_2020$Month, .keep_all = TRUE)
colnames(May_2020)[which(names(May_2020) == "May")] <- "Covid Positive Cases"
May_2020["May_2020.Month"][May_2020["May_2020.Month"]== "05"]  <- "May"
colnames(May_2020)[which(names(May_2020) == "May_2020.Year")] <- "Year"
colnames(May_2020)[which(names(May_2020) == "May_2020.Month")] <- "Month"
```

```{r}
June = sum(June_2020$Covid_positivity.numtoday)
June_2020 <- data.frame(June_2020$Month, June_2020$Year, June)
June_2020 <- June_2020 %>% distinct(June_2020$Month, .keep_all = TRUE)
colnames(June_2020)[which(names(June_2020) == "June")] <- "Covid Positive Cases"
June_2020["June_2020.Month"][June_2020["June_2020.Month"]== "06"]  <- "June"
colnames(June_2020)[which(names(June_2020) == "June_2020.Year")] <- "Year"
colnames(June_2020)[which(names(June_2020) == "June_2020.Month")] <- "Month"
```

```{r}
July = sum(July_2020$Covid_positivity.numtoday)
July_2020 <- data.frame(July_2020$Month, July_2020$Year, July)
July_2020 <- July_2020 %>% distinct(July_2020$Month, .keep_all = TRUE)
colnames(July_2020)[which(names(July_2020) == "July")] <- "Covid Positive Cases"
July_2020["July_2020.Month"][July_2020["July_2020.Month"]== "07"]  <- "July"
colnames(July_2020)[which(names(July_2020) == "July_2020.Year")] <- "Year"
colnames(July_2020)[which(names(July_2020) == "July_2020.Month")] <- "Month"
```

```{r}
August = sum(August_2020$Covid_positivity.numtoday)
August_2020 <- data.frame(August_2020$Month, August_2020$Year, August)
August_2020 <- August_2020 %>% distinct(August_2020$Month, .keep_all = TRUE)
colnames(August_2020)[which(names(August_2020) == "August")] <- "Covid Positive Cases"
August_2020["August_2020.Month"][August_2020["August_2020.Month"]== "08"]  <- "August"
colnames(August_2020)[which(names(August_2020) == "August_2020.Year")] <- "Year"
colnames(August_2020)[which(names(August_2020) == "August_2020.Month")] <- "Month"
```

```{r}
September = sum(September_2020$Covid_positivity.numtoday)
September_2020 <- data.frame(September_2020$Month, September_2020$Year, September)
September_2020 <- September_2020 %>% distinct(September_2020$Month, .keep_all = TRUE)
colnames(September_2020)[which(names(September_2020) == "September")] <- "Covid Positive Cases"
September_2020["September_2020.Month"][September_2020["September_2020.Month"]== "09"]  <- "September"
colnames(September_2020)[which(names(September_2020) == "September_2020.Year")] <- "Year"
colnames(September_2020)[which(names(September_2020) == "September_2020.Month")] <- "Month"
```

```{r}
October = sum(October_2020$Covid_positivity.numtoday)
October_2020 <- data.frame(October_2020$Month, October_2020$Year, October)
October_2020 <- October_2020 %>% distinct(October_2020$Month, .keep_all = TRUE)
colnames(October_2020)[which(names(October_2020) == "October")] <- "Covid Positive Cases"
October_2020["October_2020.Month"][October_2020["October_2020.Month"]== "10"]  <- "October"
colnames(October_2020)[which(names(October_2020) == "October_2020.Year")] <- "Year"
colnames(October_2020)[which(names(October_2020) == "October_2020.Month")] <- "Month"
```

```{r}
November = sum(November_2020$Covid_positivity.numtoday)
November_2020 <- data.frame(November_2020$Month, November_2020$Year, November)
November_2020 <- November_2020 %>% distinct(November_2020$Month, .keep_all = TRUE)
colnames(November_2020)[which(names(November_2020) == "November")] <- "Covid Positive Cases"
November_2020["November_2020.Month"][November_2020["November_2020.Month"]== "11"]  <- "November"
colnames(November_2020)[which(names(November_2020) == "November_2020.Year")] <- "Year"
colnames(November_2020)[which(names(November_2020) == "November_2020.Month")] <- "Month"
```

```{r}
December = sum(December_2020$Covid_positivity.numtoday)
December_2020 <- data.frame(December_2020$Month, December_2020$Year, December)
December_2020 <- December_2020 %>% distinct(December_2020$Month, .keep_all = TRUE)
colnames(December_2020)[which(names(December_2020) == "December")] <- "Covid Positive Cases"
December_2020["December_2020.Month"][December_2020["December_2020.Month"]== "12"]  <- "December"
colnames(December_2020)[which(names(December_2020) == "December_2020.Year")] <- "Year"
colnames(December_2020)[which(names(December_2020) == "December_2020.Month")] <- "Month"
```

```{r}
January2 = sum(January_2021$Covid_positivity.numtoday)
January_2021 <- data.frame(January_2021$Month, January_2021$Year, January2)
January_2021 <- January_2021 %>% distinct(January_2021$Month, .keep_all = TRUE)
colnames(January_2021)[which(names(January_2021) == "January2")] <- "Covid Positive Cases"
colnames(January_2021)[which(names(January_2021) == "January_2021.Year")] <- "Year"
colnames(January_2021)[which(names(January_2021) == "January_2021.Month")] <- "Month"
January_2021["Month"][January_2021["Month"]== "01"]  <- "January"
```

```{r}
February2 = sum(February_2021$Covid_positivity.numtoday)
February_2021 <- data.frame(February_2021$Month, February_2021$Year, February2)
February_2021 <- February_2021 %>% distinct(February_2021$Month, .keep_all = TRUE)
colnames(February_2021)[which(names(February_2021) == "February2")] <- "Covid Positive Cases"
colnames(February_2021)[which(names(February_2021) == "February_2021.Year")] <- "Year"
colnames(February_2021)[which(names(February_2021) == "February_2021.Month")] <- "Month"
February_2021["Month"][February_2021["Month"]== "02"]  <- "February"
```

```{r}
March2 = sum(March_2021$Covid_positivity.numtoday)
March_2021 <- data.frame(March_2021$Month, March_2021$Year, March2)
March_2021 <- March_2021 %>% distinct(March_2021$Month, .keep_all = TRUE)
colnames(March_2021)[which(names(March_2021) == "March2")] <- "Covid Positive Cases"
colnames(March_2021)[which(names(March_2021) == "March_2021.Year")] <- "Year"
colnames(March_2021)[which(names(March_2021) == "March_2021.Month")] <- "Month"
March_2021["Month"][March_2021["Month"]== "03"]  <- "March"
```

```{r}
April2 = sum(April_2021$Covid_positivity.numtoday)
April_2021 <- data.frame(April_2021$Month, April_2021$Year, April2)
April_2021 <- April_2021 %>% distinct(April_2021$Month, .keep_all = TRUE)
colnames(April_2021)[which(names(April_2021) == "April2")] <- "Covid Positive Cases"
colnames(April_2021)[which(names(April_2021) == "April_2021.Year")] <- "Year"
colnames(April_2021)[which(names(April_2021) == "April_2021.Month")] <- "Month"
April_2021["Month"][April_2021["Month"]== "04"]  <- "April"
```

```{r}
May2 = sum(May_2021$Covid_positivity.numtoday)
May_2021 <- data.frame(May_2021$Month, May_2021$Year, May2)
May_2021 <- May_2021 %>% distinct(May_2021$Month, .keep_all = TRUE)
colnames(May_2021)[which(names(May_2021) == "May2")] <- "Covid Positive Cases"
colnames(May_2021)[which(names(May_2021) == "May_2021.Year")] <- "Year"
colnames(May_2021)[which(names(May_2021) == "May_2021.Month")] <- "Month"
May_2021["Month"][May_2021["Month"]== "05"]  <- "May"
```

```{r}
June2 = sum(June_2021$Covid_positivity.numtoday)
June_2021 <- data.frame(June_2021$Month, June_2021$Year, June2)
June_2021 <- June_2021 %>% distinct(June_2021$Month, .keep_all = TRUE)
colnames(June_2021)[which(names(June_2021) == "June2")] <- "Covid Positive Cases"
colnames(June_2021)[which(names(June_2021) == "June_2021.Year")] <- "Year"
colnames(June_2021)[which(names(June_2021) == "June_2021.Month")] <- "Month"
June_2021["Month"][June_2021["Month"]== "06"]  <- "June"
```

```{r}
July2 = sum(July_2021$Covid_positivity.numtoday)
July_2021 <- data.frame(July_2021$Month, July_2021$Year, July2)
July_2021 <- July_2021 %>% distinct(July_2021$Month, .keep_all = TRUE)
colnames(July_2021)[which(names(July_2021) == "July2")] <- "Covid Positive Cases"
colnames(July_2021)[which(names(July_2021) == "July_2021.Year")] <- "Year"
colnames(July_2021)[which(names(July_2021) == "July_2021.Month")] <- "Month"
July_2021["Month"][July_2021["Month"]== "07"]  <- "July"
```

```{r}
August2 = sum(August_2021$Covid_positivity.numtoday)
August_2021 <- data.frame(August_2021$Month, August_2021$Year, August2)
August_2021 <- August_2021 %>% distinct(August_2021$Month, .keep_all = TRUE)
colnames(August_2021)[which(names(August_2021) == "August2")] <- "Covid Positive Cases"
colnames(August_2021)[which(names(August_2021) == "August_2021.Year")] <- "Year"
colnames(August_2021)[which(names(August_2021) == "August_2021.Month")] <- "Month"
August_2021["Month"][August_2021["Month"]== "08"]  <- "August"
```

```{r}
September2 = sum(September_2021$Covid_positivity.numtoday)
September_2021 <- data.frame(September_2021$Month, September_2021$Year, September2)
September_2021 <- September_2021 %>% distinct(September_2021$Month, .keep_all = TRUE)
colnames(September_2021)[which(names(September_2021) == "September2")] <- "Covid Positive Cases"
colnames(September_2021)[which(names(September_2021) == "September_2021.Year")] <- "Year"
colnames(September_2021)[which(names(September_2021) == "September_2021.Month")] <- "Month"
September_2021["Month"][September_2021["Month"]== "09"]  <- "September"
```

```{r}
October2 = sum(October_2021$Covid_positivity.numtoday)
October_2021 <- data.frame(October_2021$Month, October_2021$Year, October2)
October_2021 <- October_2021 %>% distinct(October_2021$Month, .keep_all = TRUE)
colnames(October_2021)[which(names(October_2021) == "October2")] <- "Covid Positive Cases"
colnames(October_2021)[which(names(October_2021) == "October_2021.Year")] <- "Year"
colnames(October_2021)[which(names(October_2021) == "October_2021.Month")] <- "Month"
October_2021["Month"][October_2021["Month"]== "10"]  <- "October"
```

```{r}
November2 = sum(November_2021$Covid_positivity.numtoday)
November_2021 <- data.frame(November_2021$Month, November_2021$Year, November2)
November_2021 <- November_2021 %>% distinct(November_2021$Month, .keep_all = TRUE)
colnames(November_2021)[which(names(November_2021) == "November2")] <- "Covid Positive Cases"
colnames(November_2021)[which(names(November_2021) == "November_2021.Year")] <- "Year"
colnames(November_2021)[which(names(November_2021) == "November_2021.Month")] <- "Month"
November_2021["Month"][November_2021["Month"]== "11"]  <- "November"

```

```{r}
December2 = sum(December_2021$Covid_positivity.numtoday)
December_2021 <- data.frame(December_2021$Month, December_2021$Year, December2)
December_2021 <- December_2021 %>% distinct(December_2021$Month, .keep_all = TRUE)
colnames(December_2021)[which(names(December_2021) == "December2")] <- "Covid Positive Cases"
colnames(December_2021)[which(names(December_2021) == "December_2021.Year")] <- "Year"
colnames(December_2021)[which(names(December_2021) == "December_2021.Month")] <- "Month"
December_2021["Month"][December_2021["Month"]== "12"]  <- "December"
```

View one of the dataframes for a month to ensure the data is accurate and only showing the requested information
```{r}
head(December_2021)
```


Combine all monthly data frames into a single data frame previously created as Covid_Monthly
```{r}
Covid_monthly <- rbind.data.frame(February_2020, March_2020, April_2020, May_2020, June_2020, July_2020, August_2020, September_2020, October_2020, November_2020, December_2020, January_2021, February_2021, March_2021, April_2021, May_2021, June_2021, July_2021, August_2021, September_2021, October_2021, November_2021, December_2021)

```

Export CLeaned Dataset
```{r}

write.csv(Covid_monthly,"C:/Users/Katie Schilling/Downloads/covid_monthly_clean.csv", row.names = FALSE)
```