---
title: "Granger Causality Test"
author: "Katie Schilling - 501130072"
date: "07/02/2022"
output:
  pdf_document: default
  html_document: default
---
Required tools to be loaded
```{r}
library(dplyr)
require(lmtest)

```

Import both Clean datasets
```{r}
Covid_monthly <- read.csv("C:/Users/Katie Schilling/Downloads/covid_monthly_clean.csv")
Vital_Events <- read.csv("C:/Users/Katie Schilling/Downloads/vital_events_clean.csv")

```

Combine the vital events data with the Covid Monthly data
```{r}
Final_dataset <- merge(x=Vital_Events, y=Covid_monthly, all = TRUE)
```

Check the data and ensure data merged properly
```{r}
summary(Final_dataset)
```

Change the N/A in the Covid Positive Cases to 0 so that the data is not omitted from the predictions
```{r}
Final_dataset$Covid[is.na(Final_dataset$Covid)] = 0
```

See if there are anymore NA's in the dataset
```{r}
Final_dataset %>% filter_all(any_vars(is.na(.))) 
```
Remove rows with NA as they will skew the results
```{r}
Final_dataset <- na.omit(Final_dataset)
```

Check for NA's to confirm all have been removed
```{r}
Final_dataset %>% filter_all(any_vars(is.na(.))) 
```
```{r}
summary(Final_dataset)
```

```{r}
Final_dataset$Date <- as.Date(Final_dataset$Date,"%Y-%m-%d")

```


```{r}
glimpse(Final_dataset)
```


Normalize the Data
```{r}

Final_dataset_standardized <- Final_dataset %>% mutate_each_(list(~scale(.) %>% as.vector),
                                  vars = c("Births","Marriages","Deaths","Stillbirths","Covid"))

```


```{r}
Final_ts <- ts(Final_dataset)
```


Running Granger Causality tests for each variable against each other variable one by one
Previous Granger test showed one variable against all others did not show causality, so now determining if one variable against each
other variable individually shows causality.


Covid vs. Deaths Causality
```{r}
grangertest(Covid ~ Deaths, order =12, data = Final_ts)
grangertest(Deaths ~ Covid, order =12, data = Final_ts)
```

Covid vs. Marriages Causality
```{r}
grangertest(Covid ~ Marriages, order =12, data = Final_ts)
grangertest(Marriages ~ Covid, order =12, data = Final_ts)
```

Covid vs. Stillbirths Causality
```{r}
grangertest(Covid ~ Stillbirths, order =12, data = Final_ts)
grangertest(Stillbirths ~ Covid, order =12, data = Final_ts)
```

Covid vs. Births Causality
```{r}
grangertest(Covid ~ Births, order =12, data = Final_ts)
grangertest(Births ~ Covid, order =12, data = Final_ts)
```

Deaths vs. Marriages Causality
```{r}
grangertest(Deaths ~ Marriages, order =12, data = Final_ts)
grangertest(Marriages ~ Deaths, order =12, data = Final_ts)
```

Deaths vs. Stillbirths Causality
```{r}
grangertest(Stillbirths ~ Deaths, order =12, data = Final_ts)
grangertest(Deaths ~ Stillbirths, order =12, data = Final_ts)
```

Deaths vs. Births Causality
```{r}
grangertest(Deaths ~ Births, order =12, data = Final_ts)
grangertest(Births ~ Deaths, order =12, data = Final_ts)
```

Births vs. Marriages Causality
```{r}
grangertest(Births ~ Marriages, order =12, data = Final_ts)
grangertest(Marriages ~ Births, order =12, data = Final_ts)
```

Births vs. Stillbirths Causality
```{r}
grangertest(Births ~ Stillbirths, order =12, data = Final_ts)
grangertest(Stillbirths ~ Births, order =12, data = Final_ts)
```

Marriages vs. Stillbirths Causality
```{r}
grangertest(Marriages ~ Stillbirths, order =12, data = Final_ts)
grangertest(Stillbirths ~ Marriages, order =12, data = Final_ts)
```

